Expected Dependency Pair Match:\\ Predicting translation quality with expected syntactic structure

### 2009-05-15 mtjournal.tex
Mari's last pass of changes.

% tuning target: weighted correlation with mean-removed segment HTER
\subsection{Effects of features on tuning}
% [we use weighted correlation to avoid emphasis on short segments
% (which hurts document-level and system-level utility) ]
In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
confidence interval) for the results on the two test sets (each
trained on one and tested on the other).
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.45]{tuning.eps}
  \end{center}
  \caption{Pearson's $r$ values for the two halves of the training
    set, with 95\% confidence intervals. }
  \label{fig:tuneresults}
\end{figure}
%
For each set of features, we plot the $r$ (and confidence) on both
halves of the corpus (for each half, tuned on the other half). In the
leftmost group of the figure, we include baselines: EDPM and BLEU and
TER.  Next, the tuned E feature, which (by itself) does not perform as
well as TER or EDPM.  
%
The remaining groups of three show that for each set of features, as E
(and EB) are included with others, correlation improves (though not
significantly).  These all move in the same direction. By the same
measurement, P features help beyond T -- as reported elsewhere -- but
N features do not seem to find a better correlation than TP features.
% MO suggestions:

@@

\section{Conclusion}
\label{sec:conclusion}
In this research, we explore dependency pair match (DPM) a family of
syntactically-oriented measures.  Through a corpus of human judgments,
we settle on EDPM, a member of that family with promising predictive
power.  We find that EDPM is superior to BLEU and TER in terms of
correlation with human judgments and as a per-document and
per-sentence predictor of HTER (using mean-subtraction to account for
underlying document- and segment-level difficulty).
% TO DO:  add something here regarding summary of tuning?

Several areas of future work are open to us. One obstacle to using
this method widely is the computational expense of the system. Parsing
is many times more computationally costly than word-based measures
such as BLEU or TER.  Using these syntactic techniques may be possible
as a late-pass evaluation, to identify how candidate translation
systems are performing overall.  Alternatively, the various
tree-decompositions of the DPM family could be used as system
diagnostics, by looking at relative quality of these component scores
compared to those of an alternative candidate system.

Another interesting research area is in the quality of the parser.  In
this work, we demonstrated that the
\inlinecite{charniak-johnson:2005:ACL} PCFG parser could serve in
place of the LFG parser used in
\inlinecite{owczarzak07labelleddepseval}. PCFG parsers, however, are
fairly easily adaptable to another domain. One interesting research
direction explores the intersection of genre (and genre-mismatch),
parser quality, and prediction of human performance.
% Another
% consideration is the speed of the parser: parse time currently takes
% much, much longer than BLEU, TER, or other word-based measures. 
We are interested in the trade-offs --- if any --- among parse
quality, speed, and predictive power as a translation metric.  One
possibility, for example, is to store packed forests
\cite{huang08packedforests} rather than generating an $n$-best list
only to sum across it again in calculating the expectation.

TO DO: parting words

>>>>

% tuning target: weighted correlation with mean-removed segment HTER
\subsection{Effects of features on tuning}
% [we use weighted correlation to avoid emphasis on short segments
% (which hurts document-level and system-level utility) ]
In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
confidence interval) for the results on the two test sets (each
trained on one and tested on the other).
\begin{figure}
  \begin{center}
    \includegraphics[scale=0.45]{tuning.eps}
  \end{center}
  \caption{Pearson's $r$ values for the two halves of the training
    set, with 95\% confidence intervals. }
  \label{fig:tuneresults}
\end{figure}
%
For each set of features, we plot the $r$ (and confidence) on both
halves of the corpus (for each half, tuned on the other half). In the
leftmost group of the figure, we include baselines: EDPM and BLEU and
TER (which are not tuned).  Next, the tuned E feature, which (by
itself) does not perform as well as TER or EDPM.
%
The remaining groups of three show that for each set of features, as E
(and EB) are included with others, correlation improves (though not
significantly).  These all move in the same direction. By the same
measurement, P features help beyond T -- as reported elsewhere -- but
N features do not seem to find a better correlation than TP features.
% MO suggestions:

@@

\section{Conclusion}
\label{sec:conclusion}
In this research, we explore dependency pair match (DPM) a family of
syntactically-oriented measures.  Through a corpus of human fluency
and adequacy judgments,
we settle on EDPM, a member of that family with promising predictive
power.  We find that EDPM is superior to BLEU and TER in terms of
correlation with human judgments and as a per-document and
per-sentence predictor of HTER (using mean-subtraction to account for
underlying document- and segment-level difficulty).
We experiment with including syntactic features in TERp-style tuning,
with linear weight combination, and find that including these features
improves correlation with mean-subtracted HTER, although these
improvements do not reach significance.

Several areas of future work are open to us. One obstacle to using
this method widely is the computational cost of parsing compared to
word-based measures
such as BLEU or TER.  Using these syntactic techniques may be possible
as a late-pass evaluation, to identify how candidate translation
systems are performing overall.  Alternatively, the various
tree-decompositions of the DPM family could be used as system
diagnostics, by looking at relative quality of these component scores
compared to those of an alternative candidate system.
%
Another possible approach is to store packed forests
\cite{huang08packedforests} rather than generating an $n$-best list
only to sum across it again in calculating the expectation.

Another research area relates to the quality of the parser.  In
this work, we demonstrated that the
\inlinecite{charniak-johnson:2005:ACL} PCFG parser could serve in
place of the LFG parser used in
\inlinecite{owczarzak07labelleddepseval}. PCFG parsers, however, are
fairly easily adaptable to another domain. One research
direction explores the intersection of genre (and genre-mismatch),
parser quality, and prediction of human performance.
% Another
% consideration is the speed of the parser: parse time currently takes
% much, much longer than BLEU, TER, or other word-based measures. 
We are interested in the trade-offs --- if any --- among parse
quality, genre adaptation, and predictive power as a translation metric.
